{\fontfamily{ptm}\selectfont Linear Convergence of Stochastic Nesterov's Accelerated Proximal Gradient method under Interpolation Hypothesis, Truth or Just a Dream?}

### 2025-07-18 Notebook LaTEX/SAGD Under Interpolation/main.tex
>w<, changes

\newcommand{\cov}{\ensuremath{\operatorname{Cov}}}
\newcommand{\var}{\ensuremath{\operatorname{Var}}}
\newcommand{\expect}{\ensuremath{\mathbb E}}
\usepackage{tikz}\newcommand*\circled[1]{
    \text{
        \hspace{-0.5em}\tikz[baseline=(char.base)]{
            \node[shape=circle,draw,inner sep=0.6pt](char){#1};
        }\hspace{-0.5em}
    }
}

@@

    If the label is letter, like: (a), (b), ..., then they are stated in advanced at the start of the proof and they are usually non-trivial results. 

@@

@@

@@

            Fix some $x \in \RR^n$, and suppose that an error: $w$ is made when estimating the proximal to obtain $\tilde x$ at $x$ such that it's characterized by
                w \in \partial \left[
                    z \mapsto g(z) + \langle \nabla f(x), z - x\rangle + \frac{B}{2}\Vert z - x\Vert^2
                \right](\tilde x)
            And in addition, assume that there exists some $B \ge 0$ such that $D_f(\tilde x, x) \le \frac{B}{2}\Vert \tilde x - x\Vert^2$. 
            Then, for all $z \in \RR^n$ it satisfies: 

@@

@@

            Suppose an inexact evaluation of proximal gradient operator at $x$ yield an approximation $\tilde x$ such that: 
                & (\exists w)(\exists \epsilon ): w \in \partial \left[
                    z \mapsto  g(z) + \langle \nabla f(x), z - x\rangle + \frac{B}{2}\Vert z - x\Vert^2
                \right](\tilde x), \Vert w\Vert \le \epsilon \Vert x - \tilde x\Vert. 
            Suppose that there exists some $B \ge 0$ such that $D_f(\tilde x, x) \le \frac{B}{2}\Vert \tilde x - x\Vert^2$. 

@@

@@

        \begin{lemma}[stochastic proximal gradient inequality]\label{lemma:stoch-pg-ineq}

@@

            Suppose that, $\tilde \nabla f(x)$ is a random variable which estimates $\nabla f(x)$, cause proximal gradient operator to produce error $w$, and estimate $\tilde x$ as given by \ref{lemma:inex-pg-ineq-proto}. 
            Assume there exists $B \ge 0$ be constant such that $D_f(\tilde x,x)\le B/2 \Vert x - \tilde x\Vert^2$. 
            Then, it would have $w = \nabla f(x) - \tilde \nabla f(x)$. 
            If in addition, there exists an $\epsilon \ge 0$: 

@@

                    \Vert w\Vert \Vert z - \tilde x\Vert

@@

            The reason for $w = \nabla f(x) - \tilde \nabla f(x)$ is explained in \eqref{eqn:stoch-grad-err-vec}. 

@@

                F(z) - F(\tilde x) + \frac{\max(B, L) - \mu}{2}\Vert z - x\Vert^2 + \Vert w\Vert\Vert z - \tilde x\Vert. 

@@

                    \epsilon\Vert w\Vert\Vert z - \tilde x\Vert

@@

            Otherwise, $\epsilon = 0$ is an obvious choice and, it's the smallest. 

@@

@@

            And in this case, it doesn't on $z$. 

@@

    \begin{definition}[inexact proximal gradient operator with relative error]

@@

            \item $\tilde x = \mathbf T_B^{(\epsilon)}(x | F)$ is an inexact output of proximal gradient operator by evaluating on $\tilde \nabla f(x)$. 

@@

        Then, we define the error $\epsilon$ by: 

@@

                0 & \text{else } x = \tilde x. 

@@

            \item $\epsilon$ an error which is pre-determined constant. 

@@

@@

        Here, $\epsilon$ satisfies: 

@@

                \right]} & \text{if } x \neq \tilde x \;\wedge\; z \neq \tilde x

@@

                0 & \text{else }

@@

        For this definition of the stochastic proximal gradient operator, Lemma \ref{lemma:stoch-pg-ineq} is applicable. 

@@

    % ------------------------------------------------------------------------------------------------------------------
    \begin{lemma}[properties of the sequence]\label{lemma:seq-properties}

@@

    \subsection{building up the convergence results}

@@

            Define $z_k = \alpha_k \bar x + (1 - \alpha_k)x_{k - 1}$.
            Then, the error sequence $(\epsilon_k)_{k \ge 0}$ is given by: 
                (\forall k \ge 1)\; \epsilon_k = 
                \frac{\expect_k\left[
                        \tilde \nabla f(y_k) - \nabla f(y_k)
                    \right\Vert
                    \Vert \bar x - v_{k}\Vert
                    \expect_k\left[
                        \Vert \bar x - v_{k}\Vert
                        \Vert z_k - y_k\Vert

@@

            We use Lemma \ref{lemma:stoch-pg-ineq} with $x = y_k, z = z_k, \tilde x = x_{k}$ and, $B = L$ then it means: 

@@

\subsection{Discussion}

>>>>

\newcommand{\cov}{\ensuremath{\operatorname{Cov}}}
\newcommand{\var}{\ensuremath{\operatorname{Var}}}
\newcommand{\expect}{\ensuremath{\mathbb E}}
\newcommand{\prob}{\ensuremath{\mathbb P}}
\newcommand{\dist}{\ensuremath{\operatorname{dist}}}
\usepackage{tikz}\newcommand*\circled[1]{
    \text{
        \hspace{-0.5em}\tikz[baseline=(char.base)]{
            \node[shape=circle,draw,inner sep=0.6pt](char){#1};
        }\hspace{-0.5em}
    }
}

@@

    If the label is letter, like: (a), (b), ..., then they are stated in advanced at the start of the proof and, they are usually non-trivial results. 

@@

            Note, it also has $T_\beta(x | f + g) = \hprox_{\beta^{-1}g}(x - \beta^{-1}\nabla f(x))$ in optimization literatures. 

@@

        \begin{definition}[Relative proximal gradient error ruler]\label{def:pg-err-ruler}
            Let $F$ satisfies Assumption \ref{ass:smooth-plus-nonsmooth}. 
            Fix any $x \in \RR^n$, suppose that $\tilde x$ is an estimated of $T_B(x|F)$. 
            Then the relative proximal gradient error is a set defined as
                S_B(\tilde x, x | F) := 
                \partial \left[
                    z \mapsto \partial g(x) + \langle \nabla f(x), z - x\rangle + 
                    \frac{B}{2}\Vert x - z\Vert^2
                \right](\tilde x). 
            The definition exists to simplifies the notations for the discussions. 
            When $B > 0$ by strong convexity, the set $\{z : \mathbf 0 \in S_B(z, x | F)\}$ is a singleton, conveniently. 

@@

            Fix any $x \in \RR^n$ there exists a $B \ge 0$,  $\tilde x$ be an estimate of $T_B(x | F)$ such that $D_f(\tilde x, x) \le B/2\Vert x - \tilde x\Vert^2$. 
            Let $S_B(\tilde x, x | F)$ be given by Definition \ref{def:pg-err-ruler}. 
            Then, for all $z \in \RR^n$ and, any $w \in S_B(\tilde x, x | F)$ it satisfies: 

@@

            Since $F = f + g$ satisfies Assumption \ref{ass:smooth-plus-nonsmooth}, for all $B \ge L$, it will be an obvious choice. 

@@

            Fix arbitrary $x \in \RR^n$. 
            Assume the followings: 
                \item $\tilde x$ estimates $T_B(x | F)$. 
                \item $B \ge 0$ satisfies $D_f(\tilde x, x) \le B/2\Vert x - \tilde x\Vert^2$. 
                \item Let $S_B(\tilde x, x | F)$ be given by Definition \ref{def:pg-err-ruler}. 
            Let $\epsilon \ge 0$ be such that $\Vert x - \tilde x\Vert\epsilon \ge \dist(\mathbf 0 |S_B(\tilde x, x | F))$.

@@

                + \Vert w\Vert \Vert z - \tilde x\Vert. 
            Since this is true for all $w \in S_B(\tilde x, x | F)$, it has: 
                &\le F(z) - F(\tilde x) + \frac{\max(B, L) - \mu}{2}\Vert z - x\Vert^2 
                + \dist(\mathbf 0 | S_B(\tilde x, x | F))\Vert z - \tilde x\Vert

@@

        \begin{lemma}[proximal stochastic gradient inequality]\label{lemma:pg-stoch-ineq}

@@

            Assume the following
                \item $\tilde \nabla f(x)$ is a random variable which estimates $\nabla f(x)$, which produces the estimate $\tilde x$. 
                \item There exists $B \ge 0$ such that $D_f(\tilde x,x)\le B/2 \Vert x - \tilde x\Vert^2$. 
            If in addition, there exists some $\epsilon \ge 0$: 

@@

                    \left\Vert \nabla f(x) - \tilde \nabla f(x)\right\Vert 
                    \Vert z - \tilde x\Vert

@@

            The can choose $w = \nabla f(x) - \tilde \nabla f(x) \in S_B(\tilde x, x | F)$ which is explained in \eqref{eqn:stoch-grad-err-vec}. 

@@

                F(z) - F(\tilde x) + \frac{\max(B, L) - \mu}{2}\Vert z - x\Vert^2 + \Vert w\Vert\Vert z - \tilde x\Vert
                F(z) - F(\tilde x) 
                + \frac{\max(B, L) - \mu}{2}\Vert z - x\Vert^2 
                + \left\Vert \nabla f(x) - \tilde \nabla f(x)\right\Vert\Vert z - \tilde x\Vert. 

@@

                    \epsilon\Vert x - \tilde x\Vert\Vert z - \tilde x\Vert

@@

@@

        Let's think about the edge case. 
        When $\expect [\Vert x - \tilde x\Vert\Vert z - \tilde x\Vert] = 0$, it must be that both $\Vert x - \tilde x\Vert,\Vert z - \tilde x\Vert$ are zero, indicating that $\nabla f(x) = \tilde \nabla f(x)$ and, $x = \tilde x = z$. 
        Otherwise we can assume $\prob(\Vert x - \tilde x\Vert\Vert z - \tilde x\Vert > 0) > 0$ and, the expectation has: 
            & \expect \left[
            \expect\left[
                \Vert x - \tilde x\Vert\Vert z - \tilde x\Vert \;|\; 
                \Vert x - \tilde x\Vert\Vert z - \tilde x\Vert > 0
            \right]\prob(
                \Vert x - \tilde x\Vert\Vert z - \tilde x\Vert > 0
            ). 
        Let's suppose that $\tilde \nabla f(x)$ is a random variable comes from the space: $\Omega(x)$, then it has the following: 
        {\small
            & \expect \left[
                \Vert x - \tilde x\Vert\Vert z - \tilde x\Vert \; | \; 
                \Vert x - \tilde x\Vert\Vert z - \tilde x\Vert > 0
            &\ge
            \min_{y \in \Omega(x)}\left\lbrace
                \Vert x - \tilde x\Vert
                :x\neq \tilde x = \argmin_{z}\left\lbrace
                    g(z) + \langle y, z\rangle + \frac{B}{2}\Vert z - x\Vert^2
                \right\rbrace
            \right\rbrace
            \prob(
                \Vert x - \tilde x\Vert\Vert z - \tilde x\Vert > 0
            )
            \expect\Vert z - \tilde x\Vert, 
            & \expect \left[
            \max_{y \in \Omega(x)}\left\lbrace
                \Vert \nabla f(x) - y\Vert 
            \right\rbrace\expect\Vert z - \tilde x \Vert, 
        And it would mean: 
        {\small
            \epsilon &= 
            }\\
            &\le \frac{
                \max_{y \in \Omega(x)}\Vert \nabla f(x) - y\Vert 
                \min_{y \in \Omega(x)}\left\lbrace
                    \Vert x - \tilde x\Vert:
                    x\neq \tilde x = \argmin_{z}\left\lbrace
                    g(z) + \langle y, z\rangle + \frac{B}{2}\Vert z - x\Vert^2
                \right\rbrace
                \right\rbrace
                \prob(
                    \Vert x - \tilde x\Vert\Vert z - \tilde x\Vert > 0
                )

@@

        And in this case, it has nothing to do with $z$.

@@

    \begin{definition}[proximal inexact gradient operator with relative error]

@@

            \item $\tilde x = \mathbf T_B(x | F)$ is an inexact output of proximal gradient operator by evaluating on $\tilde \nabla f(x)$. 

@@

        Then, we define the relative error $\epsilon$ by: 

@@

                \infty & \text{if } x = \tilde x, \nabla f(x) \neq \tilde \nabla f(x), 
                0 & \text{if } x = \tilde x, \nabla f(x) = \tilde \nabla f(x). 

@@

@@

            \item $\epsilon$ is an relative error, determined by $z, x, \tilde\nabla f(x)$ and, defined immediately below. 

@@

        The relative error $\epsilon$ is defined as: 

@@

                \right]} 
                & \text{if } \expect\left[\Vert x - \tilde x\Vert\Vert z - \tilde x\Vert\right] \neq 0, 
                \vspace{0.5em}

@@

                \left.\begin{cases}
                    0 & \expect \left[\left\Vert \nabla f(x) - \tilde \nabla f(x) \right\Vert \Vert z - \tilde x\Vert\right] = 0, 
                    \infty & \text{else. } 
                \end{cases}\right\rbrace
                & \text{else. }

@@

        For this definition of the stochastic proximal gradient operator, Lemma \ref{lemma:pg-stoch-ineq} is applicable. 

@@

    The following lemma defines the conditions required for the a momentum sequence to be compatible with convergence claims. 
    \begin{lemma}[relaxed momentum sequence conditions]\label{lemma:seq-properties}
        \todoinline{NOT FINISHED YET!!!}

@@

    \subsection{Building up the convergence results}

@@

@@

            We use Lemma \ref{lemma:pg-stoch-ineq} with $x = y_k, z = z_k, \tilde x = x_{k}$ and, $B = L$ then it means: 

@@

\section{Discussion, interpolation helps but it's not enough}
    \subsection{The no error case}
    \subsection{The degenerate case}
